Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-02-04 paper/main.tex
paper s5
plots

\subsubsection{Benchmark problems}
%
Experiments with real-world expensive black-box problems would be computationally infeasible. Instead, we consider classical combinatorial optimization problems as black-box optimization problems. \manuel{In this section, and following the consensus on the optimization literature, the permutations denote orderings.}{For consistency with the rest of the paper, a permutation $\sigma$ denotes a ranking also in this section, thus we use the inverse permutation $\sigma^{-1}$ in the definition of the problems. Later in the experimental section we discuss the effects of not performing this inversion.}
%
The \emph{Linear Ordering Problem (LOP)}, when minimizing, is defined as
%\footnote{For consistency with the other problems, we minimise the lower triangle instead of maximising the upper triangle.} 
\begin{equation}\label{eq:lop}
  \min_{\sigma \in S_n} f(\sigma) = \sum_{i=1}^{n} \sum_{j=1}^{i-1}  a_{\sigma^{-1}(i), \sigma^{-1}(j)}
\end{equation}
%
where $[a_{i,j}]$ is a matrix of size $n \times n$ \manuel{and $\sigma^{-1}$ is the inverse of $\sigma$}{}.

@@

During our experiments, we realized that the \emph{swap} distance used by CEGO
assumes rankings, i.e., it actually measures the \ken distance, yet CEGO did
not invert the permutation before evaluating the PFSP. We verified that this is
indeed how the implementation works and that those results match the published
ones~\citep{ZaeStoBar2014:ppsn}. Of course, the obvious question arises of what
would be the effect of inverting the permutation before calling the objective
function, so that the rankings are transformed to ordering, as done by our UMM
algorithm. Therefore, in the following, we show results of both variants, that is, the original CEGO (\CEGOorig) and our modified version that inverts the permutation before evaluation (\CEGOinv).

@@

\begin{figure*}[tb]
  \centering%
    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_IO_N-t59b11xx}
    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_SGB_N-sgb75_02}

    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_RandB_N-p40-01}
    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_RandB_N-p50-01}

    \includegraphics[width=0.49\textwidth]{../img/fitness_pfsp_rec19_txt}
    \includegraphics[width=0.49\textwidth]{../img/fitness_pfsp_rec31_txt}
\label{fig:results}
  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the PFSP instance \texttt{rec05}.\label{fig:rec05}}
\end{figure*}

>>>>

\subsubsection{Benchmark problems}
%
Experiments with real-world expensive black-box problems would be computationally infeasible. Instead, we consider classical combinatorial optimization problems as black-box optimization problems. For consistency with the rest of the paper, a permutation $\sigma$ denotes a ranking also in this section, thus we use the inverse permutation $\sigma^{-1}$ in the definition of the problems. Later in the experimental section we discuss the effects of not performing this inversion.
%
The \emph{Linear Ordering Problem (LOP)}, when minimizing, is defined as
%\footnote{For consistency with the other problems, we minimise the lower triangle instead of maximising the upper triangle.} 
\begin{equation}\label{eq:lop}
  \min_{\sigma \in S_n} f(\sigma) = \sum_{i=1}^{n} \sum_{j=1}^{i-1}  a_{\sigma^{-1}(i), \sigma^{-1}(j)}
\end{equation}
%
where $[a_{i,j}]$ is a matrix of size $n \times n$.

@@

During our experiments, we realized that the \emph{swap} distance used by CEGO
assumes rankings, i.e., it actually measures the \ken distance, yet CEGO did
not invert the permutation before evaluating the PFSP. We verified that this is
indeed how the implementation works and that those results match the published
ones~\citep{ZaeStoBar2014:ppsn}. Of course, the obvious question arises about the effect of inverting the permutation before calling the objective
function, so that the rankings are transformed to ordering, as done by our UMM
algorithm. Therefore, in the following, we show results of both variants, that is, the original CEGO (\CEGOorig) and our modified version that inverts the permutation before evaluation (\CEGOinv).

@@

\begin{figure*}[tb]
  \small
  \centering%
    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_IO_N-t59b11xx}
    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_SGB_N-sgb75_02}
\\
    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_RandB_N-p40-01}
    \begin{minipage}{0.49\linewidth}
          \texttt{RandB N-p50-01} (LOP)\\[-0.5ex]
    \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p50-01}
  \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
    \texttt{reC19} (PFSP)\\[-0.5ex]
\includegraphics[width=\textwidth]{../img/fitness_pfsp_rec19_txt}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \texttt{reC31} (PFSP)\\[-0.5ex]
    \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec31_txt}
  \end{minipage}
\label{fig:results}
  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on various instances.\label{fig:rec05}}
\end{figure*}
